---
title: "DA 301"
output: html_document
date: "2024-11-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Load libraries
library(dplyr)
library(glmnet)

#read data
data <- read.csv("new_jump_data - new_jump_data.csv")
unique(data$injury_type)

# Remove columns that are entirely NA
data <- data[, colSums(is.na(data)) < nrow(data)]

```

```{r}
model <-lm(L.R.Peak.Braking.Force ~ Peak.Propulsive.Power , data=data)
summary(model)
```


```{r}
# Example: Creating a binary variable for injury/fatigue status
data$Injured <- ifelse(data$injury_type == "NA", 0, 1)
data$Injured[is.na(data$Injured)]<- 0 

# Check distribution of the target variable
table(data$Injured)

```

There are 5043 athletes who are injured out of 10967 athletes in the dataset.

### mRSI percentage change 
```{r}
# Filter data to include only injured athletes
injured_data <- data %>% filter(Injured == 1)%>%group_by(athlete_code) %>%  # assuming athlete_id uniquely identifies each athlete
  mutate(
    mRSI_before_injury = mean(mRSI[Date < problem_date], na.rm = TRUE),  # Average mRSI before injury date
    mRSI_after_injury = mean(mRSI[Date > problem_date], na.rm = TRUE)     # Average mRSI after return date
  ) %>%
  ungroup() %>%
  
  # Calculate percentage change in mRSI
  mutate(
    mRSI_pct_change = (mRSI_after_injury / mRSI_before_injury) - 1,
    mRSI_reduction_10pct = ifelse(mRSI_pct_change <= -0.1, 1, 0)  # Binary indicator for 20% or more reduction
  )

# Filter to get athletes with a 20% or greater reduction in mRSI
significant_reduction_mRSI <- injured_data %>%
  filter(mRSI_reduction_10pct == 1)

# Plot the smoothed density graph for percentage change in mRSI
ggplot(injured_data, aes(x = mRSI_pct_change)) +
  geom_density(fill = "blue", alpha = 0.4) +
  geom_vline(xintercept = -0.1, color = "red", linetype = "dashed") +
  geom_vline(xintercept=0, color="darkgreen")+
  labs(
    title = "Density Plot of Percentage Change in mRSI for Injured Athletes",
    x = "Percentage Change in mRSI (Post / Pre Injury - 1)",
    y = "Density"
  ) +
  scale_x_continuous(expand = c(0, 0))+
  theme_minimal()
```

```{r}
# Calculate the proportion of athletes with mRSI reduction at the chosen threshold (e.g., 10%)
threshold <- 0  # Change this to your preferred threshold

# Add a binary variable for athletes below the threshold
injured_data <- injured_data %>%
  mutate(mRSI_below_threshold = ifelse(mRSI_pct_change <= threshold, 1, 0))

# Calculate proportion of athletes with mRSI below the threshold
proportion_below_threshold <- injured_data %>%
  summarise(
    total_athletes = n(),
    below_threshold = sum(mRSI_below_threshold, na.rm = TRUE),
    proportion_below_threshold = below_threshold / total_athletes
  )

# Print the results
proportion_below_threshold
```


### Peak Braking Velocity Percentage Change
```{r}
# Filter data to include only injured athletes
injured_data <- data %>% filter(Injured == 1)%>%group_by(athlete_code) %>%  # assuming athlete_id uniquely identifies each athlete
  mutate(
    braking_velocity_before_injury = mean(Peak.Braking.Velocity[Date < problem_date], na.rm = TRUE),  # Average Peak Braking Velocity before injury date
    braking_velocity_after_injury = mean(Peak.Braking.Velocity[Date > problem_date], na.rm = TRUE)     # Average mRSI after return date
  ) %>%
  ungroup() %>%
  
  # Calculate percentage change in Peak Velocity change
  mutate(
    braking_velocity_change = (braking_velocity_after_injury / braking_velocity_before_injury) - 1,
    braking_velocity_reduction_10pct = ifelse(braking_velocity_change <= -0.1, 1, 0)  # Binary indicator for 20% or more reduction
  )

# Filter to get athletes with a 20% or greater reduction in mRSI
significant_reduction_brake <- injured_data %>%
  filter(braking_velocity_reduction_10pct == 1)

# Plot the smoothed density graph for percentage change in mRSI
ggplot(injured_data, aes(x = braking_velocity_change)) +
  geom_density(fill = "purple", alpha = 0.4) +
  geom_vline(xintercept=0, color="darkgreen")+
  labs(
    title = "Percentage Change in Peak Braking Velocity for Injured Athletes",
    x = "Percentage Change in Peak Braking Velocity (Post / Pre Injury - 1)",
    y = "Density"
  ) +
  scale_x_continuous(expand = c(0, 0))+
  theme_minimal()
```

```{r}
proportion_below_threshold_velocity <- injured_data %>%
  summarise(
    total_athletes = n(),
    below_threshold_velocity = sum(braking_velocity_below_threshold, na.rm = TRUE),
    proportion_below_threshold_velocity = below_threshold_velocity / total_athletes
  )

# Print the results
proportion_below_threshold_velocity
```

### Peak Propulsive Power
```{r}
# Filter data to include only injured athletes
injured_data <- data %>% filter(Injured == 1)%>%group_by(athlete_code) %>%  # assuming athlete_id uniquely identifies each athlete
  mutate(
    propulsive_before_injury = mean(Peak.Propulsive.Force[Date < problem_date], na.rm = TRUE),  # Average Peak Braking Velocity before injury date
    propulsive_after_injury = mean(Peak.Propulsive.Force[Date > problem_date], na.rm = TRUE)     # Average mRSI after return date
  ) %>%
  ungroup() %>%
  
  # Calculate percentage change in Peak Velocity change
  mutate(
    propulsive_change = (propulsive_after_injury / propulsive_before_injury) - 1,
    propulsive_reduction_10pct = ifelse(propulsive_change <= -0.1, 1, 0)  # Binary indicator for 20% or more reduction
  )

# Filter to get athletes with a 20% or greater reduction in mRSI
significant_reduction_propulsive <- injured_data %>%
  filter(propulsive_reduction_10pct == 1)

# Plot the smoothed density graph for percentage change in mRSI
ggplot(injured_data, aes(x = propulsive_change)) +
  geom_density(fill = "orange ", alpha = 0.4) +
  geom_vline(xintercept=0, color="darkgreen")+
  labs(
    title = "Percentage Change in Peak Propulsive Force for Injured Athletes",
    x = "Percentage Change in Peak Propulsive Force (Post / Pre Injury - 1)",
    y = "Density"
  ) +
  scale_x_continuous(expand = c(0, 0))+
  theme_minimal()
```

```{r}
# Calculate the proportion of athletes with mRSI reduction at the chosen threshold (e.g., 10%)
threshold_propulsive <- 0  # Change this to your preferred threshold

# Add a binary variable for athletes below the threshold
injured_data <- injured_data %>%
  mutate(propulsive_below_threshold = ifelse(propulsive_change <= threshold_propulsive, 1, 0))

# Calculate proportion of athletes with peak braking velocity below the threshold
proportion_below_threshold_propulsive <- injured_data %>%
  summarise(
    total_athletes = n(),
    below_threshold_propulsive = sum(propulsive_below_threshold, na.rm = TRUE),
    proportion_below_threshold_propulsive= below_threshold_propulsive / total_athletes
  )

# Print the results
proportion_below_threshold_propulsive
```

### Average Braking Power
```{r}
# Filter data to include only injured athletes
injured_data <- data %>% filter(Injured == 1)%>%group_by(athlete_code) %>%  # assuming athlete_id uniquely identifies each athlete
  mutate(
    avg_brake_before_injury = mean(Avg..Braking.Power[Date < problem_date], na.rm = TRUE),  # Average Peak Braking Velocity before injury date
    avg_brake_after_injury = mean(Avg..Braking.Power[Date > problem_date], na.rm = TRUE)     # Average mRSI after return date
  ) %>%
  ungroup() %>%
  
  # Calculate percentage change in Peak Velocity change
  mutate(
    avg_brake_change = (avg_brake_after_injury / avg_brake_before_injury) - 1,
    avg_brake_reduction_10pct = ifelse(avg_brake_change <= -0.1, 1, 0)  # Binary indicator for 20% or more reduction
  )

# Filter to get athletes with a 20% or greater reduction in avg brake
significant_reduction_avg_brake <- injured_data %>%
  filter(avg_brake_reduction_10pct == 1)

# Plot the smoothed density graph for percentage change in avg brake
ggplot(injured_data, aes(x = avg_brake_change)) +
  geom_density(fill = "brown", alpha = 0.4) +
  geom_vline(xintercept=0, color="darkgreen")+
  labs(
    title = "Percentage Change in Average Braking Power for Injured Athletes",
    x = "Percentage Change in Average Braking Power (Post / Pre Injury - 1)",
    y = "Density"
  ) +
  scale_x_continuous(expand = c(0, 0))+
  theme_minimal()

```

```{r}
# Calculate the proportion of athletes with mRSI reduction at the chosen threshold (e.g., 10%)
threshold_avg_brake <- 0  # Change this to your preferred threshold

# Add a binary variable for athletes below the threshold
injured_data <- injured_data %>%
  mutate(avg_brake_below_threshold = ifelse(avg_brake_change <= threshold_avg_brake, 1, 0))

# Calculate proportion of athletes with peak braking velocity below the threshold
proportion_below_threshold_avg_brake <- injured_data %>%
  summarise(
    total_athletes = n(),
    below_threshold_avg_brake = sum(avg_brake_below_threshold, na.rm = TRUE),
    proportion_below_threshold_avg_brake= below_threshold_avg_brake / total_athletes
  )

# Print the results
proportion_below_threshold_avg_brake
```

### Peak Braking Force
```{r}
# Filter data to include only injured athletes
injured_data <- data %>% filter(Injured == 1)%>%group_by(athlete_code) %>%  # assuming athlete_id uniquely identifies each athlete
  mutate(
    braking_before_injury = mean(Peak.Braking.Force[Date < problem_date], na.rm = TRUE),  # Average Peak Braking Velocity before injury date
    braking_after_injury = mean(Peak.Braking.Force[Date > problem_date], na.rm = TRUE)     # Average mRSI after return date
  ) %>%
  ungroup() %>%
  
  # Calculate percentage change in Peak Velocity change
  mutate(
    braking_change = (braking_after_injury / braking_before_injury) - 1,
    braking_reduction_10pct = ifelse(braking_change <= -0.1, 1, 0)  # Binary indicator for 20% or more reduction
  )

# Filter to get athletes with a 20% or greater reduction in avg brake
significant_reduction_braking<- injured_data %>%
  filter(braking_reduction_10pct == 1)

# Plot the smoothed density graph for percentage change in avg brake
ggplot(injured_data, aes(x = braking_change)) +
  geom_density(fill = "navy", alpha = 0.4) +
  geom_vline(xintercept=0, color="darkgreen")+
  labs(
    title = "Percentage Change in Peak Braking Force for Injured Athletes",
    x = "Percentage Change in Peak Braking Force (Post / Pre Injury - 1)",
    y = "Density"
  ) +
  scale_x_continuous(expand = c(0, 0))+
  theme_minimal()

```

```{r}
# Calculate the proportion of athletes with mRSI reduction at the chosen threshold (e.g., 10%)
threshold_braking <- 0  # Change this to your preferred threshold

# Add a binary variable for athletes below the threshold
injured_data <- injured_data %>%
  mutate(braking_below_threshold = ifelse(braking_change <= threshold_braking, 1, 0))

# Calculate proportion of athletes with peak braking velocity below the threshold
proportion_below_threshold_braking <- injured_data %>%
  summarise(
    total_athletes = n(),
    below_threshold_braking = sum(braking_below_threshold, na.rm = TRUE),
    proportion_below_threshold_braking= below_threshold_braking / total_athletes
  )

# Print the results
proportion_below_threshold_braking
```

# mRSI Marginal Change
```{r}
mRSI_data <- data %>%
  group_by(athlete_code, Date) %>%
  summarize(
    Avg_mRSI = mean(mRSI, na.rm = TRUE),  # Calculate average mRSI per day
    .groups = "drop"  # Drop grouping after summarizing
  ) %>%
  arrange(athlete_code, Date) %>%
  mutate(
    mRSI_marginal_change = (Avg_mRSI - lag(Avg_mRSI)) / lag(Avg_mRSI)  # Day-to-day percentage change
  ) %>%
  ungroup()
```

# Visualize data for athlete BASEB1
```{r}
specific_athlete <- "BASEB1"  # Replace with desired athlete code
mRSI_data_single <- mRSI_data %>%
  filter(athlete_code == specific_athlete)  # Remove NAs
mRSI_data_single$Date <- as.Date(mRSI_data_single$Date, format = "%m-%d-%Y")

# Plot for the single athlete
ggplot(mRSI_data_single, aes(x = Date, y = mRSI_marginal_change)) +
  geom_point(color = "red") +  # Scatter plot
  geom_line(color = "black") +  # Line plot with thickness
  labs(
    title = paste("Marginal Percentage Change in mRSI for Athlete", specific_athlete),
    x = "Date",
    y = "Marginal Percentage Change in mRSI",
    caption = "Player reported having an injury on March 13. His mRSI dropped significantly on March 14."
  ) + 
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for clarity
  ) +
  scale_x_date(date_labels = "%d-%m-%Y")  # Format the date and set breaks

```

# Peak Braking Velocity Marginal % Change
```{r}
# Peak Braking Velocity Data
pbv_data <- data %>%
  group_by(athlete_code, Date) %>%  # Group by athlete_code and Date
  summarize(
    Avg_pbv = mean(Peak.Braking.Velocity, na.rm = TRUE),  # Average Peak.Braking.Velocity for each day
    .groups = "drop"  # Drop grouping after summarizing
  ) %>%
  arrange(athlete_code, Date) %>%  # Arrange by athlete_code and Date
  mutate(
    pbv_marginal_change = (Avg_pbv - lag(Avg_pbv)) / lag(Avg_pbv)  # Day-to-day percentage change
  ) %>%
  ungroup()  # Ungroup after mutation
```

# Visualize data for athlete BASEB1
```{r}
specific_athlete <- "BASEB1"  # Replace with desired athlete code
pbv_data_single <- pbv_data %>%
  filter(athlete_code == specific_athlete)  # Remove NAs
pbv_data_single$Date <- as.Date(pbv_data_single$Date, format = "%m-%d-%Y")

# Plot for the single athlete
ggplot(pbv_data_single, aes(x = Date, y = pbv_marginal_change)) +
  geom_point(color = "red") +  # Scatter plot
  geom_line(color = "black") +  # Line plot with thickness
  labs(
    title = paste("Marginal Percentage Change in Peak Braking Velocity for Athlete", specific_athlete),
    x = "Date",
    y = "Marginal Percentage Change in Peak Braking Velocity",
    caption = "Player reported having an injury on March 13."
  ) + 
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for clarity
  ) +
  scale_x_date(date_labels = "%d-%m-%Y")  # Format the date and set breaks

```

# Peak Propulsive Force
```{r}
# Peak Braking Velocity Data
ppf_data <- data %>%
  group_by(athlete_code, Date) %>%  # Group by athlete_code and Date
  summarize(
    Avg_ppf = mean(Peak.Propulsive.Force, na.rm = TRUE),  # Average Peak.Braking.Velocity for each day
    .groups = "drop"  # Drop grouping after summarizing
  ) %>%
  arrange(athlete_code, Date) %>%  # Arrange by athlete_code and Date
  mutate(
    ppf_marginal_change = (Avg_ppf - lag(Avg_ppf)) / lag(Avg_ppf)  # Day-to-day percentage change
  ) %>%
  ungroup()  # Ungroup after mutation
```

# Visualize data for athlete BASEB1
```{r}
specific_athlete <- "BASEB1"  # Replace with desired athlete code
ppf_data_single <- ppf_data %>%
  filter(athlete_code == specific_athlete)  # Remove NAs
ppf_data_single$Date <- as.Date(ppf_data_single$Date, format = "%m-%d-%Y")

# Plot for the single athlete
ggplot(ppf_data_single, aes(x = Date, y = ppf_marginal_change)) +
  geom_point(color = "red") +  # Scatter plot
  geom_line(color = "black") +  # Line plot with thickness
  labs(
    title = paste("Marginal Percentage Change in Peak Propulsive Force for Athlete", specific_athlete),
    x = "Date",
    y = "Marginal Percentage Change in Peak Propulsive Force",
    caption = "Player reported having an injury on March 13."
  ) + 
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for clarity
  ) +
  scale_x_date(date_labels = "%d-%m-%Y")  # Format the date and set breaks

```
# Avg..Braking.Power
```{r}
abp_data <- data %>%
  group_by(athlete_code, Date) %>%  # Group by athlete_code and Date
  summarize(
    Avg_abp = mean(Avg..Braking.Power, na.rm = TRUE),  # Average Peak.Braking.Velocity for each day
    .groups = "drop"  # Drop grouping after summarizing
  ) %>%
  arrange(athlete_code, Date) %>%  # Arrange by athlete_code and Date
  mutate(
    abp_marginal_change = (Avg_abp - lag(Avg_abp)) / lag(Avg_abp)  # Day-to-day percentage change
  ) %>%
  ungroup()  # Ungroup after mutation
```

# Visualize data for athlete BASEB1
```{r}
specific_athlete <- "BASEB1"  # Replace with desired athlete code
abp_data_single <- abp_data %>%
  filter(athlete_code == specific_athlete)  # Remove NAs
abp_data_single$Date <- as.Date(abp_data_single$Date, format = "%m-%d-%Y")

# Plot for the single athlete
ggplot(abp_data_single, aes(x = Date, y = abp_marginal_change)) +
  geom_point(color = "red") +  # Scatter plot
  geom_line(color = "black") +  # Line plot with thickness
  labs(
    title = paste("Marginal Percentage Change in Average Braking Power for Athlete", specific_athlete),
    x = "Date",
    y = "Marginal Percentage Change in Average Braking Power",
    caption = "Player reported having an injury on March 13."
  ) + 
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for clarity
  ) +
  scale_x_date(date_labels = "%d-%m-%Y")  # Format the date and set breaks

```

# Peak.Braking.Force
```{r}
pbf_data <- data %>%
  group_by(athlete_code, Date) %>%  # Group by athlete_code and Date
  summarize(
    Avg_pbf = mean(Peak.Braking.Force, na.rm = TRUE),  # Average Peak.Braking.Velocity for each day
    .groups = "drop"  # Drop grouping after summarizing
  ) %>%
  arrange(athlete_code, Date) %>%  # Arrange by athlete_code and Date
  mutate(
    pbf_marginal_change = (Avg_pbf - lag(Avg_pbf)) / lag(Avg_pbf)  # Day-to-day percentage change
  ) %>%
  ungroup()  # Ungroup after mutation
```

# Visualize data for athlete BASEB1
```{r}
specific_athlete <- "BASEB1"  # Replace with desired athlete code
pbf_data_single <- pbf_data %>%
  filter(athlete_code == specific_athlete)  # Remove NAs
pbf_data_single$Date <- as.Date(pbf_data_single$Date, format = "%m-%d-%Y")

ggplot(pbf_data_single, aes(x = Date, y = pbf_marginal_change)) +
  geom_point(color = "red") +  # Scatter plot
  geom_line(color = "black") +  # Line plot with thickness
  labs(
    title = paste("Marginal Percentage Change in Peak Braking Force", specific_athlete),
    x = "Date",
    y = "Marginal Percentage Change in Peak Braking Force",
    caption = "Player reported having an injury on March 13."
  ) + 
  theme_minimal() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for clarity
  ) +
  scale_x_date(date_labels = "%d-%m-%Y")  # Format the date and set breaks

```


